---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Load the data

```{r}
d0 <- read.csv("activity.csv", header=T)
```

Process the data (dates as POSIXct, POSIXt)

```{r}
library(lubridate)
d0$date <- ymd(d0$date)
str(d0)
head(d0)
```

## What is mean total number of steps taken per day?

Histogram of the total number of steps per day

```{r}
sumPerDay <- tapply(d0$steps, d0$date, sum, na.rm=T)
hist(sumPerDay, main = "Total number of steps by day", xlab = "Number of steps by day")
abline(v=mean(sumPerDay), col="red", lwd=3)
abline(v=median(sumPerDay), col="blue", lwd=3)
legend(x="topright", legend=c("mean","median"), col=c("red","blue"), bty="n", lwd=3)
summary(sumPerDay)
```

## What is the average daily activity pattern?
```{r}
averageSteps <- tapply(d0$steps, d0$interval, mean, na.rm=T)
intervalByHour <- as.numeric(names(averageSteps))/60
plot(intervalByHour, averageSteps, type = "l", ylab = "Average steps every 5 min", xlab = "Interval by hour")
```

Hour that contains the maximum number of steps

```{r}
maxAverageSteps <- which(averageSteps==max(averageSteps))
d0$interval[maxAverageSteps]/60
```

## Imputing missing values
Total number of missing values in the dataset:
```{r}
sum(is.na(d0))
```
Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r}
d1 <- d0
filler <- mean(d0$steps, na.rm = T)
d1$steps[is.na(d1$steps)] <- filler
head(d1)
```

```{r}
sumPerDayd1 <- tapply(d1$steps, d1$date, sum, na.rm=T)
hist(sumPerDayd1, main = "Total number of steps by day w/filled NA values", xlab = "Number of steps by day w/filled NA values")
abline(v=mean(sumPerDayd1), col="red", lwd=3)
abline(v=median(sumPerDayd1), col="blue", lwd=3, lty=2)
legend(x="topright", legend=c("mean","median"), col=c("red","blue"), bty="n", lwd=3)

summary(sumPerDayd1)
```

## Are there differences in activity patterns between weekdays and weekends
```{r}
weekDays <- c("Monday", "Tuesday", "Wednsday", "Thursday", "Friday")
d1$weekDay <- factor((weekdays(d1$date) %in% weekDays), levels=c(FALSE, TRUE), labels=c("weekend", "weekday"))
head(d1)
wdSubset <- subset(d1, weekDay == "weekday")
weSubset <- subset(d1, weekDay == "weekend")
weekdayMean <- tapply(wdSubset$steps, wdSubset$date, mean)
weekendMean <- tapply(weSubset$steps, weSubset$date, mean)
head(weekdayMean)
head(weekendMean)
par(mfrow = c(2,1))
plot(weekdayMean, type = "l")
plot(weekendMean, type = "l")
```
